Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-11 article/article.Rmd
Commit 2024-06-11 20:14

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OIR and the yearly
excess irradiation  for the period 1993 -- 2023, together with corresponding linear
trends. All three quantities show increasing trends, most pronounced for the
frequency of occurrence 
($`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$)
and the excess irradiation 
($`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
which are also statistically significant.
In contrast the trend of the yearly mean OIR is negligible 
($`r round(yearlyCE[var == "GLB_diff.mean", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  Despite the
very weak statistical significance of the mean OIR trend, the annual number of CE
occurrences shows a steady increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$
(Figure\nobreakspace{}\ref{fig:P-energy-N}).
Subsequently, the mean annual energy of the CE events is increasing with a rate of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$
(Figure\ \ref{fig:P-energy-sum}), witch follows the trend of the number of CE
occurrences.  We have to note that the energy related to the CE events can not be
directly linked with the total energy balance of the atmosphere. The net solar
radiation of the region is not increased, but rather redistributed through the CE
events.  This also is depicted by the "instantaneous" values of irradiance (average
over one minute), that can exceed the equivalent clear sky irradiance to a
considerable level.

@@

## Cloud enhancements climatology

Another interesting aspect of the CE cases occurrences, is their seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we have the monthly box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs on June.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
For the rest months, the occurrences gradually ramp between the maximum and minimum.
This seasonality is a combinations of the clouds occurrences, and the relevant
position of the sun during each season of the year. Unfortunately, the lack of detailed data on cloud
formation, type and location is not allowing further analysis.

>>>>

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the yearly number of
CE cases (each with duration of one minute), the yearly mean OIR and the yearly
excess irradiation  for the period 1993 -- 2023, together with corresponding linear
trends. All three quantities show increasing trends, most pronounced for the
frequency of occurrence 
($`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$)
and the excess irradiation 
($`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$),
which are also statistically significant.
In contrast the trend of the yearly mean OIR is negligible 
($`r round(yearlyCE[var == "GLB_diff.mean", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 1)`\,\text{W}/\text{m}^2/\text{year}$)
and of no statistical significance.
The average OIR for the entire period is about
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$ with standard
deviation of
$`r round(sd(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$.
The interannual variability of the data about the trend lines is quite large.
Furthermore, it tends to increase with time (at least for the quantities of panels b
and c), suggesting a significant variability in cloud patterns over the area,
possibly associated to changes in climate.

We have to note that the excess irradiation related to the CE events can not be
directly linked to the total energy balance of the atmosphere. The net solar
radiation of the region is not increased, but is rather redistributed through the CE
events.  This is also depicted by the ECE irradiance values, which can exceed the
equivalent clear sky irradiance by a significant amount.

@@

## Climatology of cloud enhancement events

Another interesting aspect of the CE events, is their distribution within the year. In
Figure\nobreakspace{}\ref{fig:relative-month-occurrences}, we have the monthly box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs on June.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
For the rest months, the occurrences gradually ramp between the maximum and minimum.
This seasonality is a combinations of the clouds occurrences, and the relevant
position of the sun during each season of the year. Unfortunately, the lack of detailed data on cloud
formation, type and location is not allowing further analysis.
